---
title: "CA3 - DAIE Project"
author: "Conor Crean & Daniel Lacey"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
---
# **Data Analytics for Immersive Environments**
### *Descriptive and Inferential Analysis of a Jungian Sandplay VR Project*

---

* #### **Table Of Contents:**

##### *Abstract:*
* *Aim and Rationale*
* *Participants and Setting*
* *Experiment Design*
* *Results Gathering*
* *Major Findings*
* *Findings/Implications*

##### *Introduction:*
* *Topic and Context*
* *Rationale*
* *Hypothesis*

##### *Method:*
* *Participants*
* *Design*
* *Materials*
* *Procedure*

##### *Results:*
* *Descriptive Statistics*
* *Inferential Statistics*
* *Statistical Tests*
* *Magnitude and Direction of Results*

##### *Discussion:*
* *Outline findings and relation to the hypothesis*
* *Limitations*

---



# Abstract


### **Aim and Rationale**

The aim of this study is to test whether a virtual Jungian sandplay environment presented to psychotherapy patients undergoing treatment for PTSD has a positive effect on their CPSS (child PTSD symptom scale) scores compared to a control group that is subjected to CBT (Cognitive Behavioural Therapy) alone.

### **Participants and Setting**

The study consists of 150 patients, aged between 18-25, age was not recorded but gender was. recorded 75 are male, 75 Female. There is an even amount across the two gender groups listed in the data. A sample of people are undergoing treatment for PTSD.

### **Experiment Design**

The sample of 150 patients were divided into three groups. There was a Control group which involved traditional CBT treatment. The second group were the Static which involved a non-animated model content in VR. The third group was the Animated group which involved animated model content within VR. 

### **Results Gathering**

The results were gathered from the three distinct groups over 12 weeks (about 3 months) during 50-minute treatment sessions. The PTSD results were measured using observer-rated and self-reported mechanisms. The measurements were taken at the start and end of the study  

### **Major Findings**


### **Findings/Implications**

If this experiment shows a similar positive therapeutic effect to non-VR Jungian Sandplay then it opens the possibility of easy access to the Jungian Sandplay therapeutic method via VR instead of having to attend a physical location to receive the therapy. Thus, VR could possibly fill the role as a useful analogue to real life Jungian Sandplay making it a much more widely available therapeutic method to alleviate PTSD.

# Introduction

### **Topic and Context**

Does a virtual Jungian sandplay environment have a meaningful effect on CPSS scores in comparison to CBT.  

### **Rationale**


### **Hypothesis**


# Method

### **Participants**

150 young adults were recruited for the study (M=75, F=75) using random sampling. They were divided into 3 groups:  

* Control (traditional CBT with no VR) 

* Static (non-animated model content with VR) 

* Animated (animated model content with VR) 

All the patients were young adults between 18 – 25 years old. Gender information was recorded. All groups were of equal size. Each group was evenly split by sex (M=25, F=25). 

### **Design**

The 3 groups underwent 12 weeks (about 3 months) of treatment for 50 minutes per week with a therapist. During that time, the participants either underwent traditional CBT or used one of the 2 VR apps (Static non-animated model content, Animated model content). PTSD measurements for all groups were measured using the observer-rated and the CPSS self-report mechanisms. Measurements were taken pre-trial and post-trial for each group.

### **Materials**

### **Procedure**


# Results

##### *Cleaning data*
```{r , echo=FALSE}
library(readr)
daie_ca3_data_10 <- read_csv("daie_ca3_data_10.csv")
#View(daie_ca3_data_10)
head(daie_ca3_data_10,10)
summary(daie_ca3_data_10)
tail(daie_ca3_data_10,5)
#install.packages("tidyverse")
library(tidyverse)
data()
glimpse(daie_ca3_data_10)
class(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$test_group)
#Replacing the incorrect value
daie_ca3_data_10$test_group <- str_replace(daie_ca3_data_10$test_group,"Cotrol", "Control")
#Checking the Group values
unique(daie_ca3_data_10$test_group)
#Checking for null values in the data
is.null(daie_ca3_data_10)
is.null(daie_ca3_data_10$pre_trial_cpss)
is.null(daie_ca3_data_10$post_trial_cpss)
is.null(daie_ca3_data_10$pre_trial_or)
is.null(daie_ca3_data_10$post_trial_or)
is.na(daie_ca3_data_10)
daie_ca3_data_10 <-daie_ca3_data_10 %>% drop_na()
is.na(daie_ca3_data_10)
#Checked for negative values
sum(daie_ca3_data_10 < 0)
#Checked for values greater than 10
#Data for the entire set due to the first heading in the data 
sum(daie_ca3_data_10 > 10)
sum(daie_ca3_data_10$pre_trial_cpss > 10)
sum(daie_ca3_data_10$post_trial_cpss > 10)
sum(daie_ca3_data_10$pre_trial_or > 10)
sum(daie_ca3_data_10$post_trial_or > 10)
```


### **Descriptive Statistics**

```{r, echo=TRUE}
#Show 10 of the first few values of the data set
head(daie_ca3_data_10,10)
#Show last 10 values of data
tail(daie_ca3_data_10,5)
#Gives an overall summary of the data 
summary(daie_ca3_data_10)
daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower = min(pre_trial_cpss),
          Average = mean(pre_trial_cpss),
          Upper = max(pre_trial_cpss),
          Difference = max(pre_trial_cpss)-min(pre_trial_cpss))
  daie_ca3_data_10%>%
  drop_na(test_group)%>%  
  group_by(test_group)%>%
  summarise(Lower1 = min(post_trial_cpss),
          Average1 = mean(post_trial_cpss),
          Upper1 = max(post_trial_cpss),
          Difference1 = max(post_trial_cpss)-min(post_trial_cpss))
  
  daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower2 = min(pre_trial_or),
          Average2 = mean(pre_trial_or),
          Upper2 = max(pre_trial_or),
          Difference2 = max(pre_trial_or)-min(pre_trial_or))
  
  daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower3 = min(post_trial_or),
          Average3 = mean(post_trial_or),
          Upper3 = max(post_trial_or),
          Difference2 = max(post_trial_or)-min(post_trial_or))
static_data<-c(6.0914,5.4188,6.1032,5.4124)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(static_data,xlab="Static VR Group",ylab="Trial Scores Mean", main="Static Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Orange", border="Blue", ylim=c(0,10))
animated_data<-c(5.998163265,5.387755102,6.037959184,5.364897959)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(animated_data,xlab="Animated VR Group",ylab="Trial Scores Mean", main="Animated Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Red", border="Blue", ylim=c(0,10))
control_data<-c(6.2172,5.323,6.2638,5.3314)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(control_data,xlab="Control Group",ylab="Trial Scores Mean", main="Control Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Green", border="Blue", ylim=c(0,10))
```




### **Inferential Statistics**

---

#### *Histograms showing the normal distribution between the four Pre and Post assesments*



```{r , echo=FALSE}
boxplot(pre_trial_cpss~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups", col="Orange",
        ylab="Pre Trial CPSS")
boxplot(post_trial_cpss~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups",col="Red",
        ylab="Post Trial CPSS")
boxplot(pre_trial_or~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups",col="Green",
        ylab="Pre Trial OR")
boxplot(post_trial_or~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",col="Blue",
        xlab="Test Groups",
        ylab="Post Trial OR")
#Histogram for Static Group 
df2<-subset(daie_ca3_data_10, test_group!="Control" & test_group!="Animated")
hist(df2$pre_trial_cpss,
col="orange3",
main = "Pre Trial CPSS - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$post_trial_cpss,
col="orange1",
main = "Post Trial CPSS - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$pre_trial_or,
col="orange3",
main = "Pre Trial OR - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$post_trial_or,
col="orange1",
main = "Post Trial OR - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,20))
#Histogram for Animated Group 
df3<-subset(daie_ca3_data_10, test_group!="Static" & test_group!="Control")
hist(df3$pre_trial_cpss,
col="red3",
main = "Pre Trial CPSS - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$post_trial_cpss,
col="red",
main = "Post Trial CPSS - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$pre_trial_or,
col="red3",
main = "Pre Trial OR - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$post_trial_or,
col="red",
main = "Post Trial OR - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
#Histogram for Control Group 
df1<-subset(daie_ca3_data_10, test_group!="Static" & test_group!="Animated")
df1
  
hist(df1$pre_trial_cpss,
col="green4",
main = "Pre Trial CPSS - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,10))
hist(df1$post_trial_cpss,
col="green",
main = "Post Trial CPSS - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,20))
hist(df1$pre_trial_or,
col="green4",
main = "Pre Trial OR - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df1$post_trial_or,
col="green",
main = "Post Trial OR - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
#Normal distribution of overall trial scores
hist(daie_ca3_data_10$pre_trial_cpss,
main = "Pre Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 10)
hist(daie_ca3_data_10$post_trial_cpss,
main = "Post Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 10)
hist(daie_ca3_data_10$pre_trial_or,
main = "Pre Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 10)
hist(daie_ca3_data_10$post_trial_or,
main = "Post Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 10)
```

* #### *QQ Plots*

```{r QQ Plot - daie_ca3_data_10, echo=FALSE}
#q plot of Control Group
qqnorm(df1$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df1$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df1$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df1$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df1$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df1$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df1$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df1$post_trial_or,
       col = "darkorchid", lwd = 1)
#q plot of Static Group
qqnorm(df2$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df2$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df2$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df2$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df2$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df2$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df2$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df2$post_trial_or,
       col = "darkorchid", lwd = 1)
#q plot of Animated Group
qqnorm(df3$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df3$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df3$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df3$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df3$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df3$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df3$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df3$post_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$post_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_or,
       col = "darkorchid", lwd = 1)
```


### **Statistical Tests**
```{r}
summary(df1)
summary(df2)
summary(df3)
sample.mean <- mean(daie_ca3_data_10$pre_trial_cpss)
print(sample.mean)

  
sample.n <- length(daie_ca3_data_10$pre_trial_cpss)
sample.sd <- sd(daie_ca3_data_10$pre_trial_cpss)
sample.se <- sample.sd/sqrt(sample.n)
print(sample.se)
  
alpha = 0.05
degrees.freedom = sample.n - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
print(t.score)
  
margin.error <- t.score * sample.se
  
lower.bound <- sample.mean - margin.error
upper.bound <- sample.mean + margin.error
print(c(lower.bound,upper.bound))
  
  
```



### **Magnitude and Direction of Results**

# Discussion

### **Outline findings and relation to the hypothesis**
### **Limitations**



# References

*
*
*
*