---
title: "CA3 - DAIE Project"
author: "Conor Crean & Daniel Lacey"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
# **Data Analytics for Immersive Environments**
### *Descriptive and Inferential Analysis of a Jungian Sandplay VR Project*

---

* #### **Table Of Contents:**

##### *Abstract:*
* *Aim and Rationale*
* *Participants and Setting*
* *Experiment Design*
* *Results Gathering*
* *Major Findings*
* *Findings/Implications*

##### *Introduction:*
* *Topic and Context*
* *Rationale*
* *Hypothesis*

##### *Method:*
* *Participants*
* *Design*
* *Materials*
* *Procedure*

##### *Results:*
* *Descriptive Statistics*
* *Inferential Statistics*
* *Statistical Tests*
* *Magnitude and Direction of Results*

##### *Discussion:*
* *Outline findings and relation to the hypothesis*
* *Limitations*

---



# Abstract


### **Aim and Rationale**

The aim of this study is to test whether a virtual Jungian sandplay environment presented to psychotherapy patients undergoing treatment for PTSD has a positive effect on their CPSS (child PTSD symptom scale) scores compared to a control group that is subjected to CBT (Cognitive Behavioural Therapy) alone.

### **Participants and Setting**

The study consists of 150 patients, aged between 18-25, age was not recorded but gender was. recorded 75 are male, 75 Female. There is an even amount across the two gender groups listed in the data. A sample of people are undergoing treatment for PTSD.

### **Experiment Design**

The sample of 150 patients were divided into three groups. There was a Control group which involved traditional CBT treatment. The second group were the Static which involved a non-animated model content in VR. The third group was the Animated group which involved animated model content within VR. 

### **Results Gathering**

The results were gathered from the three distinct groups over 12 weeks (about 3 months) during 50-minute treatment sessions. The PTSD results were measured using observer-rated and self-reported mechanisms. The measurements were taken at the start and end of the study  

### **Major Findings**


### **Findings/Implications**

If this experiment shows a similar positive therapeutic effect to non-VR Jungian Sandplay then it opens the possibility of easy access to the Jungian Sandplay therapeutic method via VR instead of having to attend a physical location to receive the therapy. Thus, VR could possibly fill the role as a useful analogue to real life Jungian Sandplay making it a much more widely available therapeutic method to alleviate PTSD.

# Introduction

### **Topic and Context**

Does a virtual Jungian Sandplay environment have a meaningful effect on CPSS scores in comparison to CBT.  

### **Rationale**

The rationale behind this study is to ascertain the possible effectiveness of VR based therapies and their potential role in augmenting or complementing established psychotherapies.

### **Hypothesis**

We plan to investigate the effect of VR based therapy. In this experiment, we will test the effectiveness of VR augmented Jungian Sandplay therapy. We expect that there is likely a significant use case for VR based therapy in a modern therapeutic environment.

# Method

### **Participants**

150 young adults were recruited for the study (M=75, F=75) using random sampling. They were divided into 3 groups:  

* Control (traditional CBT with no VR) 

* Static (non-animated model content with VR) 

* Animated (animated model content with VR) 

All the patients were young adults between 18 – 25 years old. Gender information was recorded. All groups were of equal size. Each group was evenly split by sex (M=25, F=25). 

### **Design**

The 3 groups underwent 12 weeks (about 3 months) of treatment for 50 minutes per week with a therapist. During that time, the participants either underwent traditional CBT or used one of the 2 VR apps (Static non-animated model content, Animated model content). PTSD measurements for all groups were measured using the observer-rated and the CPSS self-report mechanisms. Measurements were taken pre-trial and post-trial for each group.

### **Materials**

The groups using the VR app used a tablet with the app installed. They were also given a pen and board to self-report their level of symptoms. 

The control group just used a pen and board to self-report their symptoms whilst attending a CBT session with a therapist.

### **Procedure**

1. Each group of patients in the static, animated and control groups self-report their PTSD symptoms using the Child PTSD Symptom Scale (CPSS) before their respective therapy sessions. The attending therapists will also record their patient observations pre-therapy. 

2. The static and animated groups will then use their respective VR apps on the tablets provided for their therapy sessions. The control group will attend ordinary CBT sessions. 

3. All groups will then self-report their symptoms using the Child PTSD Symptom Scale (CPSS) post-therapy on the boards provided. The attending therapists will also record their observations post-therapy.


# Results

##### *Cleaning data*
Before the data set could be properly analyzed first it needed to be cleaned to remove any errors or missing data points. 

The tidyverse package was installed, tidyverse is a collection of open-source packages used for data import, tidying, manipulation, and data visualization. 

```{r , echo=FALSE}
library(readr)
daie_ca3_data_10 <- read_csv("daie_ca3_data_10.csv")
#View(daie_ca3_data_10)

summary(daie_ca3_data_10)

#install.packages("tidyverse")
library(tidyverse)
data()
```

The next step involved looking at the different headings in the data set (gender, test_group ,etc.).  

Then the headings were checked for unique values. From the data set the test_group headings displayed four unique variables instead of the three expected. A heading labelled “Cotrol” was the fourth unexpected variable, this was clearly a mistake in the spelling of Control.  

This was then changed using the str_replace function which switched the string with the correct spelling. 


```{r , echo=FALSE}
glimpse(daie_ca3_data_10)
class(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$test_group)
#Replacing the incorrect value
daie_ca3_data_10$test_group <- str_replace(daie_ca3_data_10$test_group,"Cotrol", "Control")
#Checking the Group values
unique(daie_ca3_data_10$test_group)
```

The next step was checking for rows with missing values and removing those rows from the data to avoid errors in the results. From the summary function we see that there is one missing value in the pre_trail_or group. 

```{r , echo=FALSE}
#Checking for null values in the data

is.null(daie_ca3_data_10$pre_trial_or)

```


The entire data set was checked by using the is.na function on the excel file. That printed a long list False results only showing one True statement in row 144 under the pre_trail_or heading meaning that is where the missing value was located. 

 

To remove the row the following statement was used. 
daie_ca3_data_10 <-daie_ca3_data_10 %>% drop_na()

```{r , echo=FALSE}
daie_ca3_data_10 <-daie_ca3_data_10 %>% drop_na()

#Checked for negative values
sum(daie_ca3_data_10 < 0)
#Checked for values greater than 10
#Data for the entire set due to the first heading in the data 
sum(daie_ca3_data_10 > 10)
sum(daie_ca3_data_10$pre_trial_cpss > 10)
sum(daie_ca3_data_10$post_trial_cpss > 10)
sum(daie_ca3_data_10$pre_trial_or > 10)
sum(daie_ca3_data_10$post_trial_or > 10)
```


### **Descriptive Statistics**

Below is the summary of the results from the patient's data. From an initial look at the statistical analysis that the mean output for pre trail Observer-Rated (OR) and pre trail Child PTSD Symptom Scale Self Report Version had higher values than the post trail results. This is for the overall group; further analysis is needed to find the values for each group.  


```{r, echo=TRUE}
#Show 10 of the first few values of the data set
head(daie_ca3_data_10,10)
#Show last 10 values of data
tail(daie_ca3_data_10,5)
#Gives an overall summary of the data 
summary(daie_ca3_data_10)
daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower = min(pre_trial_cpss),
          Average = mean(pre_trial_cpss),
          Upper = max(pre_trial_cpss),
          Difference = max(pre_trial_cpss)-min(pre_trial_cpss))
  daie_ca3_data_10%>%
  drop_na(test_group)%>%  
  group_by(test_group)%>%
  summarise(Lower1 = min(post_trial_cpss),
          Average1 = mean(post_trial_cpss),
          Upper1 = max(post_trial_cpss),
          Difference1 = max(post_trial_cpss)-min(post_trial_cpss))
  
  daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower2 = min(pre_trial_or),
          Average2 = mean(pre_trial_or),
          Upper2 = max(pre_trial_or),
          Difference2 = max(pre_trial_or)-min(pre_trial_or))
  
  daie_ca3_data_10%>%
  drop_na(test_group)%>% 
  group_by(test_group)%>%
  summarise(Lower3 = min(post_trial_or),
          Average3 = mean(post_trial_or),
          Upper3 = max(post_trial_or),
          Difference2 = max(post_trial_or)-min(post_trial_or))
static_data<-c(6.0914,5.4188,6.1032,5.4124)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(static_data,xlab="Static VR Group",ylab="Trial Scores Mean", main="Static Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Orange", border="Blue", ylim=c(0,10))
animated_data<-c(5.998163265,5.387755102,6.037959184,5.364897959)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(animated_data,xlab="Animated VR Group",ylab="Trial Scores Mean", main="Animated Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Red", border="Blue", ylim=c(0,10))
control_data<-c(6.2172,5.323,6.2638,5.3314)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(control_data,xlab="Control Group",ylab="Trial Scores Mean", main="Control Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Green", border="Blue", ylim=c(0,10))
```




### **Inferential Statistics**

---

#### *Histograms showing the normal distribution between the four Pre and Post assessments*


* #### *Box plots of the three groups Pre and Post trials*
```{r , echo=FALSE}


boxplot(pre_trial_cpss~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups", col="Orange",
        ylab="Pre Trial CPSS")


boxplot(post_trial_cpss~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups",col="Red",
        ylab="Post Trial CPSS")


boxplot(pre_trial_or~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",
        xlab="Test Groups",col="Green",
        ylab="Pre Trial OR")


boxplot(post_trial_or~test_group,data=daie_ca3_data_10,
        main="Jungian Sandplay",col="Blue",
        xlab="Test Groups",
        ylab="Post Trial OR")
```


```{r , echo=FALSE}
#Histogram for Static Group 
df2<-subset(daie_ca3_data_10, test_group!="Control" & test_group!="Animated")
hist(df2$pre_trial_cpss,
col="orange3",
main = "Pre Trial CPSS - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$post_trial_cpss,
col="orange1",
main = "Post Trial CPSS - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$pre_trial_or,
col="orange3",
main = "Pre Trial OR - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df2$post_trial_or,
col="orange1",
main = "Post Trial OR - Static Group",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,20))
#Histogram for Animated Group 
df3<-subset(daie_ca3_data_10, test_group!="Static" & test_group!="Control")
hist(df3$pre_trial_cpss,
col="red3",
main = "Pre Trial CPSS - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$post_trial_cpss,
col="red",
main = "Post Trial CPSS - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$pre_trial_or,
col="red3",


main = "Pre Trial OR - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df3$post_trial_or,
col="red",




main = "Post Trial OR - Animated",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
#Histogram for Control Group 
df1<-subset(daie_ca3_data_10, test_group!="Static" & test_group!="Animated")

  
hist(df1$pre_trial_cpss,
col="green4",
main = "Pre Trial CPSS - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,10))
hist(df1$post_trial_cpss,
col="green",



main = "Post Trial CPSS - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,20))
hist(df1$pre_trial_or,
col="green4",



main = "Pre Trial OR - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))
hist(df1$post_trial_or,
col="green",



main = "Post Trial OR - Control",
xlab = "Score",
ylab = "Frequency",
breaks = 10,
xlim = c(0,10),
ylim = c(0,15))




```


* ### *Normal distribution of overall trial scores*
To check for normal distribution in the population a histogram was created for the test groups row of numbers. To create this the following function was used.   

This was applied to the 4 stages of the assessments used on the participants. 
```{r , echo=FALSE}

hist(daie_ca3_data_10$pre_trial_cpss,
main = "Pre Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 10)


hist(daie_ca3_data_10$post_trial_cpss,
main = "Post Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 10)


hist(daie_ca3_data_10$pre_trial_or,
main = "Pre Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 10)



hist(daie_ca3_data_10$post_trial_or,
main = "Post Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 10)
```

The data was shown to be normally distributed across the 4 trials

* ### *QQ Plots Groups and Trial type*

```{r QQ Plot - Groups, echo=FALSE}
#q plot of Control Group
qqnorm(df1$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df1$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df1$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df1$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df1$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df1$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df1$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df1$post_trial_or,
       col = "darkorchid", lwd = 1)
#q plot of Static Group
qqnorm(df2$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df2$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df2$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df2$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df2$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df2$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df2$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df2$post_trial_or,
       col = "darkorchid", lwd = 1)
#q plot of Animated Group
qqnorm(df3$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df3$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df3$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(df3$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(df3$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(df3$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(df3$post_trial_or,
       pch = 2, frame = FALSE)
qqline(df3$post_trial_or,
       col = "darkorchid", lwd = 1)
```

The data looks to be normally distributed across the QQ plots 
* ### *QQ Plots Trials*

QQ plot was created, the plots show that the data is roughly along the straight line of the graph

*QQ Plot - Pre Trial CPSS*
```{r QQ Plot - daie_ca3_data_10, echo=FALSE}


qqnorm(daie_ca3_data_10$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
```

*QQ Plot - Post Trial CPSS*
```{r QQ Plot - daie_ca3_data_10 Post CPSS, echo=FALSE}
qqnorm(daie_ca3_data_10$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_cpss,
       col = "darkorchid", lwd = 1)
```

*QQ Plot - Pre Trial OR*
```{r QQ Plot - daie_ca3_data_10 Pre OR, echo=FALSE}
qqnorm(daie_ca3_data_10$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_or,
       col = "darkorchid", lwd = 1)
```

*QQ Plot - Post Trial OR*
```{r QQ Plot - daie_ca3_data_10 Post OR, echo=FALSE}
qqnorm(daie_ca3_data_10$post_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_or,
       col = "darkorchid", lwd = 1)
```


### **Statistical Tests**


*Confidence Interval for Pre Trial CPSS*
```{r, echo=FALSE}

sample.mean <- mean(daie_ca3_data_10$pre_trial_cpss)
print(sample.mean)


sample.n <- length(daie_ca3_data_10$pre_trial_cpss)
sample.sd <- sd(daie_ca3_data_10$pre_trial_cpss)
sample.se <- sample.sd/sqrt(sample.n)
print(sample.se)
  
alpha = 0.05
degrees.freedom = sample.n - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
print(t.score)
  
margin.error <- t.score * sample.se
  
lower.bound <- sample.mean - margin.error
upper.bound <- sample.mean + margin.error
print(c(lower.bound,upper.bound))
  


```

*Confidence Interval for Post Trial CPSS*

```{r, echo=FALSE}
sample.n <- length(daie_ca3_data_10$post_trial_cpss)
sample.sd <- sd(daie_ca3_data_10$post_trial_cpss)
sample.se <- sample.sd/sqrt(sample.n)
print(sample.se)
  
alpha = 0.05
degrees.freedom = sample.n - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
print(t.score)
  
margin.error <- t.score * sample.se
  
lower.bound <- sample.mean - margin.error
upper.bound <- sample.mean + margin.error
print(c(lower.bound,upper.bound))

```

*Confidence Interval for Pre Trial OR*
```{r, echo=FALSE}
sample.n <- length(daie_ca3_data_10$pre_trial_or)
sample.sd <- sd(daie_ca3_data_10$pre_trial_or)
sample.se <- sample.sd/sqrt(sample.n)
print(sample.se)
  
alpha = 0.05
degrees.freedom = sample.n - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
print(t.score)
  
margin.error <- t.score * sample.se
  
lower.bound <- sample.mean - margin.error
upper.bound <- sample.mean + margin.error
print(c(lower.bound,upper.bound))




```

*Confidence Interval for Post Trial OR*
```{r, echo=FALSE}
sample.n <- length(daie_ca3_data_10$post_trial_or)
sample.sd <- sd(daie_ca3_data_10$post_trial_or)
sample.se <- sample.sd/sqrt(sample.n)
print(sample.se)
  
alpha = 0.05
degrees.freedom = sample.n - 1
t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
print(t.score)
  
margin.error <- t.score * sample.se
  
lower.bound <- sample.mean - margin.error
upper.bound <- sample.mean + margin.error
print(c(lower.bound,upper.bound))



```



### **Magnitude and Direction of Results**

# Discussion

---

### **Outline findings and relation to the hypothesis**


The experiment appears to show that there is a measurable positive effect in the use of VR augmented therapy.


### **Limitations**

We found no confounding variables within the study.


# References


---

* Clean your data with R. R programming for beginners. https://www.youtube.com/watch?v=sV5lwAJ7vnQ. 

* NNK (2022). R – Replace String with Another String or Character. https://sparkbyexamples.com/r-programming/replace-string-with-another-string-in-r/. 

* Mcleod, S. (2019). How to Write a Lab Report | Simply Psychology. [online] Simplypsychology.org. https://www.simplypsychology.org/research-report.html. 

* How to make Bar chart in R | RStudio |. https://www.youtube.com/watch?v=bzTY6_8fK7U 

* NNK (2022). How to Remove Rows with NA in R. [online]. https://sparkbyexamples.com/r-programming/remove-rows-with-na-in-r/#:~:text=By%20using%20na.