---
title: "CA3 - DAIE Project"
author: "Conor Crean & Daniel Lacey"
date: "`r Sys.Date()`"
output: html_document
---
# **Data Analytics for Immersive Environments**
## **Descriptive and Inferential Analysis of a Jungian Sandplay VR Project**

### **Table Of Contents**

#### Abstract
* *Aim and Rationale*
* *Participants and Setting*
* *Experiment Design*
* *Results Gathering*
* *Major Findings*
* *Findings/Implications*

#### Introduction
* *Topic and Context*
* *Rationale*
* *Hypothesis*

#### Method
* *Participants*
* *Design*
* *Materials*
* *Procedure*

#### Results
* *Descriptive Statistics*
* *Inferential Statistics*
* *Statistical Tests*
* *Magnitude and Direction of Results*

#### Discussion

* *Outline findings and relation to the hypothesis*
* *Limitations*


### Abstract

The aim of this study is to test whether a virtual Jungian sandplay environment presented to psychotherapy patients undergoing treatment for PTSD has a positive effect on their CPSS (child PTSD symptom scale) scores compared to a control group that is subjected to CBT (Cognitive Behavioural Therapy) alone.





# Descriptive Statistics

```{r}

library(readr)
daie_ca3_data_10 <- read_csv("daie_ca3_data_10.csv")
#View(daie_ca3_data_10)

head(daie_ca3_data_10,10)

summary(daie_ca3_data_10)

tail(daie_ca3_data_10,5)

static_data<-c(6.0914,5.4188,6.1032,5.4124)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")

barplot(static_data,xlab="Static VR Group",ylab="Trial Scores Mean", main="Static Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Orange", border="Blue", ylim=c(0,10))


animated_data<-c(6.0462,5.315,6.0736,5.3418)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")

barplot(animated_data,xlab="Animated VR Group",ylab="Trial Scores Mean", main="Animated Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Red", border="Blue", ylim=c(0,10))


control_data<-c(6.2172,5.323,6.2638,5.3314)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")

barplot(control_data,xlab="Control Group",ylab="Trial Scores Mean", main="Control Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Green", border="Blue", ylim=c(0,10))

```



# Cleaning data
```{r}

#install.packages("tidyverse")
library(tidyverse)
glimpse(daie_ca3_data_10)
class(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$test_group)


#Replacing the incorrect value
daie_ca3_data_10$test_group <- str_replace(daie_ca3_data_10$test_group,"Cotrol", "Control")

#Checking the Group values
unique(daie_ca3_data_10$test_group)

#Checking for null values in the data

is.null(daie_ca3_data_10)

is.null(daie_ca3_data_10$pre_trial_cpss)

is.null(daie_ca3_data_10$post_trial_cpss)

is.null(daie_ca3_data_10$pre_trial_or)

is.null(daie_ca3_data_10$post_trial_or)


#Checked for negative values
sum(daie_ca3_data_10 < 0)

#Checked for values greater than 10
sum(daie_ca3_data_10 > 10)

```





#Inferenial Statistics


##Histograms showing the normal distribution between the 4 Pre and Post assesments

```{r}

cdata <- subset(daie_ca3_data_10, test_group == 0)
sdata <- subset(daie_ca3_data_10, test_group == 1)
adata <- subset(daie_ca3_data_10, test_group == 2)

cpre_cpssmean <- mean(cdata$pre_trial_cpss)
cpre_cpsssd <- sd(cdata$pre_trial_cpss)



hist(daie_ca3_data_10$post_trial_cpss,
main = "Post Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 8)


hist(daie_ca3_data_10$pre_trial_or,
main = "Pre Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 8)


hist(daie_ca3_data_10$post_trial_or,
main = "Post Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 8)



```

## QQ Plots

```{r QQ Plot - daie_ca3_data_10 }

qqnorm(daie_ca3_data_10$pre_trial_cpss,
       pch = 2, frame = FALSE)

qqline(daie_ca3_data_10$pre_trial_cpss,
       col = "darkorchid", lwd = 1)



qqnorm(daie_ca3_data_10$post_trial_cpss,
       pch = 2, frame = FALSE)

qqline(daie_ca3_data_10$post_trial_cpss,
       col = "darkorchid", lwd = 1)




qqnorm(daie_ca3_data_10$pre_trial_or,
       pch = 2, frame = FALSE)

qqline(daie_ca3_data_10$pre_trial_or,
       col = "darkorchid", lwd = 1)



qqnorm(daie_ca3_data_10$post_trial_or,
       pch = 2, frame = FALSE)

qqline(daie_ca3_data_10$post_trial_or,
       col = "darkorchid", lwd = 1)




```

```{r}







```

