---
title: "CA3 - DAIE Project"
author: "Conor Crean & Daniel Lacey"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
---
# **Data Analytics for Immersive Environments**
### *Descriptive and Inferential Analysis of a Jungian Sandplay VR Project*

---

* #### **Table Of Contents:**

##### *Abstract:*
* *Aim and Rationale*
* *Participants and Setting*
* *Experiment Design*
* *Results Gathering*
* *Major Findings*
* *Findings/Implications*

##### *Introduction:*
* *Topic and Context*
* *Rationale*
* *Hypothesis*

##### *Method:*
* *Participants*
* *Design*
* *Materials*
* *Procedure*

##### *Results:*
* *Descriptive Statistics*
* *Inferential Statistics*
* *Statistical Tests*
* *Magnitude and Direction of Results*

##### *Discussion:*
* *Outline findings and relation to the hypothesis*
* *Limitations*

---

## Abstract

---

#### **Aim and Rationale**

The aim of this study is to test whether a virtual Jungian sandplay environment presented to psychotherapy patients undergoing treatment for PTSD has a positive effect on their CPSS (child PTSD symptom scale) scores compared to a control group that is subjected to CBT (Cognitive Behavioural Therapy) alone.

#### **Participants and Setting**

The study consists of 150 patients, aged between 18-25, age was not recorded but gender was. recorded 75 are male, 75 Female. There is an even amount across the two gender groups listed in the data. A sample of people are undergoing treatment for PTSD.

#### **Experiment Design**

The sample of 150 patients were divided into three groups. There was a Control group which involved traditional CBT treatment. The second group were the Static which involved a non-animated model content in VR. The third group was the Animated group which involved animated model content within VR. 

#### **Results Gathering**

The results were gathered from the three distinct groups over 12 weeks (about 3 months) during 50-minute treatment sessions. The PTSD results were measured using observer-rated and self-reported mechanisms. The measurements were taken at the start and end of the study  

#### **Major Findings**


#### **Findings/Implications**

If this experiment shows a similar positive therapeutic effect to non-VR Jungian Sandplay then it opens the possibility of easy access to the Jungian Sandplay therapeutic method via VR instead of having to attend a physical location to receive the therapy. Thus, VR could possibly fill the role as a useful analogue to real life Jungian Sandplay making it a much more widely available therapeutic method to alleviate PTSD.

## Introduction

---

#### **Topic and Context**

Does a virtual Jungian sandplay environment have a meaningful effect on CPSS scores in comparison to CBT.  

#### **Rationale**


#### **Hypothesis**


## Method

---

#### **Participants**

150 young adults were recruited for the study (M=75, F=75) using random sampling. They were divided into 3 groups:  

* Control (traditional CBT with no VR) 

* Static (non-animated model content with VR) 

* Animated (animated model content with VR) 

All the patients were young adults between 18 – 25 years old. Gender information was recorded. All groups were of equal size. Each group was evenly split by sex (M=25, F=25). 

#### **Design**

The 3 groups underwent 12 weeks (about 3 months) of treatment for 50 minutes per week with a therapist. During that time, the participants either underwent traditional CBT or used one of the 2 VR apps (Static non-animated model content, Animated model content). PTSD measurements for all groups were measured using the observer-rated and the CPSS self-report mechanisms. Measurements were taken pre-trial and post-trial for each group.

#### **Materials**

#### **Procedure**


## Results

---
#### **Descriptive Statistics**

```{r}
library(readr)
daie_ca3_data_10 <- read_csv("daie_ca3_data_10.csv")
#View(daie_ca3_data_10)
head(daie_ca3_data_10,10)
summary(daie_ca3_data_10)
tail(daie_ca3_data_10,5)
static_data<-c(6.0914,5.4188,6.1032,5.4124)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(static_data,xlab="Static VR Group",ylab="Trial Scores Mean", main="Static Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Orange", border="Blue", ylim=c(0,10))
animated_data<-c(6.0462,5.315,6.0736,5.3418)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(animated_data,xlab="Animated VR Group",ylab="Trial Scores Mean", main="Animated Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Red", border="Blue", ylim=c(0,10))
control_data<-c(6.2172,5.323,6.2638,5.3314)
CPSSOR<-c("Pre Trial CPSS","Post Trial CPSS","Pre Trial OR","Post Trial OR")
barplot(control_data,xlab="Control Group",ylab="Trial Scores Mean", main="Control Group - Pre and Post Trial CPSS and OR Scores", names.arg=CPSSOR, col="Green", border="Blue", ylim=c(0,10))
```



##### *Cleaning data*
```{r}
#install.packages("tidyverse")
library(tidyverse)
glimpse(daie_ca3_data_10)
class(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$gender)
unique(daie_ca3_data_10$test_group)
#Replacing the incorrect value
daie_ca3_data_10$test_group <- str_replace(daie_ca3_data_10$test_group,"Cotrol", "Control")
#Checking the Group values
unique(daie_ca3_data_10$test_group)
#Checking for null values in the data
is.null(daie_ca3_data_10)
is.null(daie_ca3_data_10$pre_trial_cpss)
is.null(daie_ca3_data_10$post_trial_cpss)
is.null(daie_ca3_data_10$pre_trial_or)
is.null(daie_ca3_data_10$post_trial_or)
#Checked for negative values
sum(daie_ca3_data_10 < 0)
#Checked for values greater than 10
sum(daie_ca3_data_10 > 10)
```





#### **Inferential Statistics**


* #### *Histograms showing the normal distribution between the 4 Pre and Post assesments*

```{r}
cdata <- subset(daie_ca3_data_10, test_group == 0)
sdata <- subset(daie_ca3_data_10, test_group == 1)
adata <- subset(daie_ca3_data_10, test_group == 2)
cpre_cpssmean <- mean(cdata$pre_trial_cpss)
cpre_cpsssd <- sd(cdata$pre_trial_cpss)
hist(daie_ca3_data_10$post_trial_cpss,
main = "Post Trial CPSS",
xlab = "Score",
ylab = "Frequency",
breaks = 8)
hist(daie_ca3_data_10$pre_trial_or,
main = "Pre Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 8)
hist(daie_ca3_data_10$post_trial_or,
main = "Post Trial OR",
xlab = "Score",
ylab = "Frequency",
breaks = 8)
```

* #### *QQ Plots*

```{r QQ Plot - daie_ca3_data_10 }
qqnorm(daie_ca3_data_10$pre_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$post_trial_cpss,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_cpss,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$pre_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$pre_trial_or,
       col = "darkorchid", lwd = 1)
qqnorm(daie_ca3_data_10$post_trial_or,
       pch = 2, frame = FALSE)
qqline(daie_ca3_data_10$post_trial_or,
       col = "darkorchid", lwd = 1)
```

```{r}
```



#### **Statistical Tests**

#### **Magnitude and Direction of Results**

## Discussion

#### **Outline findings and relation to the hypothesis**
#### **Limitations**



## References

*
*
*
*